---
title: "P8105_hw5_hm2947"
author: "Matthew Ma"
date: "2022-11-16"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(viridis)
library(patchwork)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 1

The code chunk below imports the data in individual spreadsheets contained in `./data/zip_data/`. To do this, I create a dataframe that includes the list of all files in that directory and the complete path to each file. As a next step, I `map` over paths and import data using the `read_csv` function. Finally, I `unnest` the result of `map`.

```{r}
full_df = 
  tibble(
    files = list.files("data/zip_data/"),
    path = str_c("data/zip_data/", files)
  ) %>% 
  mutate(data = map(path, read_csv)) %>% 
  unnest()
```

The result of the previous code chunk isn't tidy -- data are wide rather than long, and some important variables are included as parts of others. The code chunk below tides the data using string manipulations on the file, converting from wide to long, and selecting relevant variables. 

```{r}
tidy_df = 
  full_df %>% 
  mutate(
    files = str_replace(files, ".csv", ""),
    group = str_sub(files, 1, 3)) %>% 
  pivot_longer(
    week_1:week_8,
    names_to = "week",
    values_to = "outcome",
    names_prefix = "week_") %>% 
  mutate(week = as.numeric(week)) %>% 
  select(group, subj = files, week, outcome)
```

Finally, the code chunk below creates a plot showing individual data, faceted by group. 

```{r}
tidy_df %>% 
  ggplot(aes(x = week, y = outcome, group = subj, color = group)) + 
  geom_point() + 
  geom_path() + 
  facet_grid(~group)
```

This plot suggests high within-subject correlation -- subjects who start above average end up above average, and those that start below average end up below average. Subjects in the control group generally don't change over time, but those in the experiment group increase their outcome in a roughly linear way. 

## Problem 2

### Data cleaning and tidying process

```{r, message=FALSE, warning=FALSE}
homicide_raw <- read_csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv") %>% 
  janitor::clean_names() %>%
  mutate(
    state = toupper(state),
    city_state = str_c(city, state, sep = ",") %>% 
    str_replace("Tulsa,AL","Tulsa,OK"))
homicide_raw
    
homicide_tidy = 
  homicide_raw %>% 
  mutate(
    status = ifelse(disposition != "Closed by arrest", "unsolved", "solved")) %>% 
  group_by(city_state) %>% 
  summarise(n_unsolved = sum(status == "unsolved"),
            n_total = n())

homicide_tidy

```

There are `r nrow(homicide_raw)` observations of `r ncol(homicide_raw)` variables regarding the homicide in 50 cities from different states across the U.S. Variables include homicides' information including sex, race, age, and location of the crimes. By creating `city_state` variable, there are some variances among the variable where some of the state name are not fully capitalized and that the city Tulsa are misplaced into the state "AL" instead of "OK". We fix it in the code chunk above.

### Baltimore prop test estimate

We create a function `myproptest` instead of setting up x and y separately to estimate the proportion of Baltimore, MD.

```{r}

baltimore = 
  homicide_tidy %>% 
  filter(city_state == "Baltimore,MD")

pt_f = function(n_unsolved, n_total) {
  prop.test(n_unsolved, n_total, p = 0.5, alternative = "two.sided") %>%
    broom::tidy() %>%
    select(estimate, conf.low, conf.high)}

baltimore_prop =
  pt_f(pull(baltimore, n_unsolved), pull(baltimore, n_total)) %>%
  rename(lower_limit = conf.low,
    upper_limit = conf.high)
baltimore_prop
```

### Every city prop test estimate


```{r}
homicide_prop = 
  homicide_tidy %>% 
  mutate(prop_homicide = map2(n_unsolved, n_total, ~ prop.test(.x, .y) %>%
      broom::tidy())) %>%
  unnest(prop_homicide) %>%
  select(city_state, estimate, conf.low, conf.high) %>%
  rename(
    lower_limit = conf.low,
    upper_limit = conf.high)

homicide_prop
```

The resulting dataframe `homicide_prop` contains `r nrow(homicide_prop)` results of all 50 cities which contains the estimates and confidence intervals.

### Creating a plot to show the estimates and CIs for each city

```{r}
homicide_prop %>%
  mutate(
    city_state = fct_reorder(city_state, estimate, .desc = TRUE)) %>% 
  ggplot(aes(x = city_state, y = estimate, color = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower_limit, ymax = upper_limit, width = .3)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") +
  labs(
    x = "Location (City/State)",
    y = "Proportion of unsolved homicides in 50 cities acorss the U.S. (2007-2017)"
  )
  
```


## Problem 3

### Simulation t-test

```{r}
sim_t_test = function(n = 30, mu, sigma = 5) {
  sim_data = tibble(
    x = rnorm( n = 30, mean = mu, sd = sigma)
  )
  sim_data %>% 
    t.test(mu = 0, conf.level = 0.95, alternative = c("two.sided")) %>%
    broom::tidy() %>% 
    select(estimate, p.value)
}

set.seed(1)
sim_df0 = 
  expand_grid(
    mu = 0,
    iterate = 1:5000) %>% 
  mutate(ttest_df = map(.x = mu, ~sim_t_test(mu = .x))) %>% 
  unnest(ttest_df)

sim_df0

sim_df = 
  expand_grid(
    mu = c(1:6),
    iterate = 1:5000) %>% 
  mutate(ttest_df = map(.x = mu, ~sim_t_test(mu = .x))) %>% 
  unnest(ttest_df) %>% 
  rename(p_value = p.value)


sim_df

```


### Making Plots

```{r}
sim_df %>% 
  mutate(
    reject = if_else(p_value < 0.05, "T", "F")) %>%
  group_by(mu, reject) %>%
  summarize(
    num_reject = n()) %>%
  mutate(
    prop_reject = num_reject/5000) %>%
  filter(reject == "T") %>%
  ggplot(
    aes(x = mu, y = prop_reject)) +
  geom_point() + 
  geom_line() + 
  labs(
    title = "Association between effect size and power of the t-test",
    x = "True Mean",
    y = "Proportion of reject null"
  )

```

According to the plot, as the true mean increases from 1 to 6, the proportion of rejected null also increase from approximately 0.2 to 1 following a rapid increase from x = 1 to x = 3. Then a slow increase until it has reached the maximum 1. The larger the effect size, the greater the power. 

### Making two plots of average estimate comparing all data and those with null rejected.

```{r}

plot_1 = 
  sim_df %>% 
  group_by(mu) %>% 
  summarize(
    avg_est = mean(estimate)) %>% 
  ggplot(aes(x = mu, y = avg_est)) + 
  geom_point() +
  geom_line() + 
  labs( 
    title = "Average estimate (all)",
    x = "True mean",
    y = "Average Estimate")

plot_2 = 
  sim_df %>% 
  mutate(
    reject = if_else(p_value < 0.05, "T", "F")) %>%
  filter(reject == "T") %>% 
  group_by(mu) %>% 
  summarize(avg_est = mean(estimate)) %>% 
  ggplot(aes(x = mu, y = avg_est)) + 
  geom_point() +
  geom_line() + 
  labs( 
    title = "Average estimate (null rejected)",
    x = "True mean",
    y = "Average Estimate")

plot_1 + plot_2

```

For the rejected null samples the sample average is barely equal to the true mean when the true mean is smaller and equal to 3. The sample average is higher than the true mean when the true mean increases pass 3. The reason for this is because the probability of rejecting the null hypothesis increases with increased effect size. As samples with higher true means are simulated, more effects are applied.

